Variable Selection in Convex Function Estimation

### 2013-10-02 tex/scam/appendix.tex
formatted optimization section
added lots of proof details

Take the derivative with respect to the value $f_k(x_k)$ and set it equal to zero, we get that
\begin{align*}
\int_{\mathbf{x}_{-k}} p(\mathbf{x}) f_k(x_k) d \mathbf{x}_{-k} &= 
  \int_{\mathbf{x}_{-k}} p(\mathbf{x}) 
      ( f(\mathbf{x})-\sum_{k' \neq k} f^*_{k'}(x_{k'})-\mu^*) d \mathbf{x}_{-k} \\  
p(x_k) f_k(x_k) &= p(x_k) \int_{\mathbf{x}_{-k}} 
       p(\mathbf{x}_{-k}) ( f(\mathbf{x}) - \sum_{k' \neq k} f^*_{k'} (x_{k'})-\mu^*) 
              d \mathbf{x}_{-k} \\
f_k(x_k) &= \int_{\mathbf{x}_{-k}} 
       p(\mathbf{x}_{-k}) (f(\mathbf{x}) -\mu^*) d \mathbf{x}_{-k} 
 \end{align*}
Where the second line follows because $p(\mathbf{x}) = \prod_k p(x_k)$ by product distribution assumption. The third line follows because $\int_{\mathbf{x}_{-k}} p(\mathbf{x}_{-k}) f^*_{k'}(x_{k'}) d\mathbf{x}_{-k} = 0$ for all $k' \neq k$ by the constraint that $\E[f^*_{k'}(X_{k'})] = 0$.

@@

 The next proposition, combined with the previous Lemma, proves Theorem~\ref{thm:convex_faithful}.

@@

For a function $g : \R^s \rightarrow \R$, define $\hat{R}_s(g) \coloneqq \| f_0 + w - g \|_n^2$ as the objective of the \emph{restricted} regression and define $R_s(g) \coloneqq \E | f_0(X)  + w - g(X) |^2$ as the population risk.

@@

\subsubsection{Proof}

\begin{lemma}
Suppose assumptions A1' and A4 hold.

@@

\begin{align*}
R(f^{*(s-1)}) - R(f^{*s}) &= \E(f^{*(s-1)}(X) - f_0(X))^2 - \E(f^{*s}(X) - f_0(X))^2 \\
    &= \E( f^{*(s-1)}(X) - f^{*s}(X) + f^{*s}(X) - f_0(X))^2 - \E(f^{*s}(X) - f_0(X))^2 \\
    &= \E( f^{*(s-1)}(X) - f^{*s}(X))^2 - 2 \E[ (f^{*(s-1)}(X) - f^{*s}(X)) (f^{*s}(X) - f_0(X) ) ]
\end{align*}

@@

Suppose $\hat{f}$ has at most $s-1$ non-zero components. Then
\begin{align*}
\hat{R}_s( \hat{f}) &\geq R_s(\hat{f}) - \tau_n \\
	&\geq R_s(f^{*(s-1)}) - \tau_n \\
	&\geq R_s(f^{*s}) + \alpha - \tau_n \\
	&\geq \hat{R}_s(f^{*s}) + \alpha - 2 \tau_n \\
	&\geq \hat{R}_s(f^{*s} - \bar{f}^{*s}) -\tau'_n + \alpha - 2\tau_n 
\end{align*}
Where $\tau_n$ is the deviation between empirical risk and true risk and $\tau'_n$ is the approximation error incurred by empirically sampling $f^{*s}$.

@@

For a fixed $f'$, we have that, by definition
\[
\| f_0 + w - f' \|_n^2 = \| f_0 - f'\|_n^2 + 2 \langle w, f_0 - f'\rangle_n + \| w \|_n^2 \\
\]
Because $f_0(X^{(i)}) - f'^(X^{(i)})$ is bounded by $2sB$, $\| f_0 - f'\|_n^2$ is the empirical average of $n$ random variables bounded by $4(sB)^2$.

@@

>>>>

Take the derivative with respect to the value $f_k(x_k)$ and set it equal to zero, we get that
\begin{align*}
\int_{\mathbf{x}_{-k}} p(\mathbf{x}) f_k(x_k) d \mathbf{x}_{-k} &= 
  \int_{\mathbf{x}_{-k}} p(\mathbf{x}) 
      ( f(\mathbf{x})-\sum_{k' \neq k} f^*_{k'}(x_{k'})-\mu^*) d \mathbf{x}_{-k} \\  
p(x_k) f_k(x_k) &= \int_{\mathbf{x}_{-k}} p(x_k)
     p(\mathbf{x}_{-k}) ( f(\mathbf{x}) - \sum_{k' \neq k} f^*_{k'} (x_{k'})-\mu^*) 
              d \mathbf{x}_{-k} \\
f_k(x_k) &= \int_{\mathbf{x}_{-k}} 
       p(\mathbf{x}_{-k}) (f(\mathbf{x}) -\mu^*) d \mathbf{x}_{-k} 
 \end{align*}
Where the second line follows because $p(\mathbf{x}) = \prod_k p(x_k)$ by product distribution assumption. The third line follows because $\int_{\mathbf{x}_{-k}} p(\mathbf{x}_{-k}) f^*_{k'}(x_{k'}) d\mathbf{x}_{-k} = 0$ for all $k' \neq k$ by the constraint that $\E[f^*_{k'}(X_{k'})] = 0$.

@@

Our reasoning actually proves something stronger. It proves the following lemma:

\begin{lemma}
\label{lem:additive_component}

$f^*_k(x_k) = \E[ f(X) | x_k] - \mu^*$ is the unique optimal solution to 
\[
min_{f_k} \{ \E( f(X) - \E f(X) - \sum_{k'\neq k} f_{k'}(X_{k'}) - f_k (X_k))^2 \,:\, \E f_k(X_k) = 0 \}
\]
for \emph{any} set of component functions $\{ f_{k'} \}_{k'\neq k}$ that satisfies $\E f_{k'}(X_{k'}) = 0$.
\end{lemma}

 The next proposition, combined with Lemma~\ref{lem:int_reduction}, proves Theorem~\ref{thm:convex_faithful}.

@@

For a function $g : \R^s \rightarrow \R$, define $\hat{R}_s(g) \coloneqq \| f_0 + w - \bar{f_0 + w} - g \|_n^2$ as the objective of the \emph{restricted} regression and define $R_s(g) \coloneqq \E | f_0(X)  + w - \mu - g(X) |^2$ as the population risk, where $\bar{f_0 + w} = \frac{1}{n} \sum_i f_0(X^{(i)}) + w_i $ and $\mu = \E f_0(X)$. Note that we \emph{subtract} out the mean in the two risk definitions.

@@

\subsubsection{Proof}
\label{sec:false_negative_proof_proof}

\begin{lemma}
\label{lem:minus_one_risk_increase}
Suppose assumptions A1' and A4 hold.

@@

\begin{align*}
R(f^{*(s-1)}) - R(f^{*s}) &= \E\Big(f^{*(s-1)}(X) - f_0(X) + \mu\Big)^2 
                              - \E\Big(f^{*s}(X) - f_0(X) + \mu\Big)^2 \\
    &= \E\Big( f^{*(s-1)}(X) - f^{*s}(X) + f^{*s}(X) - f_0(X) + \mu\Big)^2 
                                 - \E\Big(f^{*s}(X) - f_0(X) + \mu\Big)^2 \\
    &= \E\Big( f^{*(s-1)}(X) - f^{*s}(X)\Big)^2 
                - 2 \E\Big[ (f^{*(s-1)}(X) - f^{*s}(X)) (f^{*s}(X) - (f_0(X)-\mu)) \Big]
\end{align*}

We will argue that all the components of the additive function $f^{*(s-)}$ are also in $f^{*s}$. Let us denote the components of $f^{*s} = \sum_{k=1}^s f_k^*$. We will now invoke Lemma~\ref{lem:additive_component}, which is valid because we assume $X_1,...,X_s$ are independent by assumption A1'. By Lemma~\ref{lem:additive_component}, if we set $f^*_k = 0$, the resulting additive function $\sum_{k'\neq k} f_{k'}^*$ minimizes the population risk subject to the constraint that $f_k = 0$. By definition, $f^{*s}$ is $\argmin_k \sum_{k'\neq k} f_{k'}^*$ and thus share components with $f^{*s}$.

Therefore, there exist some $k$ such that $f^{*(s-1)} - f^{*s} = f^*_k$, and we can continue the bound
\begin{align*}
R(f^{*(s-1)}) - R(f^{*s}) &= \E f^*_k(X_k)^2 - 2 \E[ f^*_k(X_k) ( f^{*s}(X) - (f_0(X)-\mu) )] \\
    &= \E f^*_k(X_k)^2 - 2 \E[ f^*_k(X_k) f^{*s} (X)] + 2 \E[ f^*_k(X_k) (f_0(X)-\mu) ] 
\end{align*}

\begin{align*}
\E[ f^*_k(X_k) f^{*s} (X)]  &= \E\Big[ f^*_k (X_k) \E[ f^{*s} (X) \,|\, X_k ] \Big] = \E f^*_k(X_k)^2  \\
\E[ f^*_k(X_k)(f_0(X) - \mu) ] &= \E\Big[ f^*_k(X_k) \E[ f_0(X) - \mu \,|\, X_k ] \Big] = \E f^*_k(X_k)^2 
\end{align*}

Where we used the fact that $\E f^*_{k'} (X_{k'}) = 0$ for all $k'$ and the fact that $\E[ f_0(X) - \mu \,|\, X_k] = f^*_k(X_k)$ (Lemma~\ref{lem:int_reduction}).

Thus, $R(f^{*(s-1)}) - R(f^{*s}) \geq \E f^*_k(X_k)^2 \geq \alpha$ by Assumption A4.

@@

Now that we bounded the cost of approximating $f^{*s}$ with the empirically centered $f^{*s} - \bar{f}^{*s}$, we move on to the proof of the main result.

Suppose $\hat{f}$ has at most $s-1$ non-zero components. Then
\begin{align*}
\hat{R}_s( \hat{f}) &\geq R_s(\hat{f}) - \tau_n \\
	&\geq R_s(f^{*(s-1)}) - \tau_n \\
	&\geq R_s(f^{*s}) + \alpha - \tau_n \\
	&\geq \hat{R}_s(f^{*s}) + \alpha - 2 \tau_n \\
	&\geq \hat{R}_s(f^{*s} - \bar{f}^{*s}) -\tau'_n + \alpha - 2\tau_n 
\end{align*}
The third line follows from Lemma~\ref{lem:minus_one_risk_increase}.
$\tau_n$ is the deviation between empirical risk and true risk and $\tau'_n$ is the approximation error incurred by empirically sampling $f^{*s}$.

@@

For a fixed $f'$, we have that, by definition
\[
\| f_0 + w - f' \|_n^2 = \| f_0 - f'\|_n^2 + 2 \langle w, f_0 - f'\rangle_n + \| w \|_n^2 \\
\]

Because $f_0(X^{(i)}) - f'^(X^{(i)})$ is bounded by $2sB$, $\| f_0 - f'\|_n^2$ is the empirical average of $n$ random variables bounded by $4(sB)^2$.

@@

% DO NOT CHANGE; RefTex variables -minx

%%% Local Variables: ***
%%% mode:latex ***
%%% TeX-master: "icml14/scam_icml.tex" ***
%%% End: ***
